---
title: "Pharmacy-Managed Warfarin Dosing Service Analysis for Fiscal Year 2018"
author: "Brian Gulbis, PharmD, BCPS"
date: "October 2018"
output:
  tufte::tufte_handout:
    latex_engine: xelatex
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(
    tidy = FALSE, 
    echo = FALSE, 
    message = FALSE, 
    warning = FALSE, 
    cache.extra = packageVersion('tufte'), 
    kable.force.latex = TRUE
)

options(htmltools.dir.version = FALSE)
```

```{r data, message=FALSE}
library(tidyverse)
library(lubridate)
library(broom)
library(themebg)

x <- dirr::get_rds("../data/tidy/fy2018")

start_date <- mdy("7/1/2017", tz = "US/Central")
end_date <- mdy("6/30/2018", tz = "US/Central")

data_ts_2017 <- read_rds("../data/tidy/fy2017/data_timeseries_location.Rds")

# ttr_pharm_prev <- data_warfarin %>%
#     filter(!is.na(perc.time),
#            warfarin_start >= mdy("7/1/2016", tz = "US/Central"),
#            warfarin_start <= mdy("6/30/2017", tz = "US/Central"),
#            group == "pharmacy") %>%
#     mutate_at("perc.time", as.numeric) %>%
#     mutate_at("perc.time", funs(. * 100)) %>%
#     mutate_at("group", str_replace_all, pattern = "pharmacy", replacement = "pharmacy_fy16")

ttr_pharm_curr <- data_warfarin %>%
    filter(
        !is.na(perc.time),
        warfarin_start >= start_date,
        warfarin_start <= end_date,
        group == "pharmacy"
    ) %>%
    mutate_at("perc.time", as.numeric) %>%
    mutate_at("perc.time", funs(. * 100)) 

ttr_trad_curr <- data_warfarin %>%
    filter(
        !is.na(perc.time),
        warfarin_start >= start_date,
        warfarin_start <= end_date,
        group == "traditional"
    ) %>%
    mutate_at("perc.time", as.numeric) %>%
    mutate_at("perc.time", funs(. * 100)) 

my_scale_fill <- function(..., values) {
    scale_fill_manual(
        "", 
        values = c("#969696", "#636363"), 
        labels = c("Traditional", "Pharmacy")
    )
}

```

# Summary

Utilization of the Pharmacy-Managed Warfarin Dosing Service continued to increase in FY17. The areas with the highest utilization were Rehab, CVIMU, and Stroke units. Patients managed by pharmacy had a greater time in therapeutic range and were less likely to have an unplanned return to a MH facility within 30 days. 

# Service Utilization

There has been a 20% increase in the number of warfarin patients being managed by the Pharmacy Dosing Service over the past two fiscal years (see Figure 1). Conversely, there has been a 25% decrease in patients with traditional warfarin management (warfarin managed by a physician or advanced practice provider) during this time. 

```{r, fig.cap="Number of Warfarin patients managed by Pharmacy Dosing Service compared with traditional management on a weekly basis from July 2016 through June 2018."}
data_timeseries_location %>%
    bind_rows(data_ts_2017) %>%
    filter(fy == 2017 | fy == 2018) %>%
    group_by(week) %>%
    summarize_at(c("pharmacy", "traditional"), sum, na.rm = TRUE) %>%
    gather(group, num, pharmacy, traditional) %>%
    ggplot(aes(x = week, y = num)) +
    geom_line(aes(linetype = group)) +
    geom_smooth(aes(linetype = group), method = "lm", color = "black", se = FALSE) +
    # ggtitle("Warfarin Volume") +
    scale_x_datetime(NULL, date_labels = "%b %Y") +
    scale_y_continuous("Weekly Patients", breaks = seq(0, 300, 50)) +
    scale_linetype_discrete(NULL, labels = c("Pharmacy", "Traditional")) +
    theme_bg(base_family = "serif") +
    theme(legend.position = "bottom")
```

The highest volume of warfarin use occured in the Heart Failure IMU, followed by CIMU and CVIMU (see Figure 2). Among the units with the highest warfarin volume, the Pharmacy Dosing Service was responsible for managing warfarin in 50% or more of the patients in the CVIMU, Rehab, 3-Cullen, and Stroke units. Utilization of the Pharmacy Dosing Service remains low in both HFIMU and CIMU, although Clinical Pharmacists and Clinical Pharmacist-Specialists participated in the management of warfarin in both of these areas.

```{r, fig.align='center', fig.height=4, out.height='3in', fig.cap="Top 10 hospital units administering warfarin in FY18."}
# For comparison, FY16 data is provided: the X indicates warfarin managed by Pharmacy Dosing Service in each unit in FY16, the | indicates the total warfarin in each unit in FY16.


data_timeseries_location %>%
    filter(fy == 2018) %>%
    group_by(order.location) %>%
    summarize_at(c("pharmacy", "traditional"), sum, na.rm = TRUE) %>%
    mutate(total = pharmacy + traditional) %>%
    arrange(desc(total)) %>%
    top_n(10, total) %>%
    gather(group, num, pharmacy, traditional) %>%
    mutate_at("group", factor, levels = c("traditional", "pharmacy")) %>%
    mutate_at("order.location", as_factor) %>%
    mutate_at("order.location", fct_rev) %>%
    ggplot(aes(x = order.location, y = num, fill = group)) +
    geom_bar(stat = "identity") +
    # geom_errorbar(aes(x = order.location, ymin = total, ymax = total), data = units16, width = 0.5, color = "grey35") +
    # geom_point(aes(x = order.location, y = pharmacy), data = units16, shape = 4, color = "grey35") +
    xlab("Hospital Unit") +
    ylab("Warfarin Doses Administered") +
    # scale_fill_brewer(NULL, palette = "Greys", labels = c("Traditional", "Pharmacy")) +
    my_scale_fill() +
    coord_flip() +
    theme_bg(base_family = "serif", yticks = FALSE) +
    theme(legend.position = "bottom")
```

The Pharmacy Dosing Service managed a higher percent of patients with venous thromboembolism and mechanical heart valves, while the traditional management was responsible for a higher percent of patients with mechanical circulatory support devices (see Figure 3). There was a similar percent of patients with atrial fibrillation in both groups.

```{r, fig.align='center', fig.height=4, out.height='3in', fig.cap="Percent of patients with each warfarin indication among patients managed by pharmacy and traditional."}
data_warfarin %>%
    filter(
        warfarin_start >= start_date,
        warfarin_start <= end_date
    ) %>%
    group_by(group) %>%
    summarize_if(is.logical, funs(mean(., na.rm = TRUE) * 100)) %>%
    gather(indication, val, afib:other) %>%
    arrange(group, desc(val)) %>%
    mutate_at("group", factor, levels = c("traditional", "pharmacy")) %>%
    mutate_at("indication", fct_inorder) %>%
    mutate_at("indication", fct_rev) %>%
    filter(val > 0) %>%
    ggplot(aes(x = indication, y = val, fill = group)) +
    geom_bar(stat = "identity", position = "dodge") +
    scale_x_discrete(
        "Indication", 
        labels = c(
            "VAD", 
            "Hypercoaguable", 
            "Stroke", 
            "Thrombus", 
            "Valve", 
            "PE", 
            "DVT", 
            "A.fib"
        )
    ) +
    ylab("Patients (%)") +
    my_scale_fill() +
    coord_flip() +
    theme_bg(base_family = "serif", yticks = FALSE) +
    theme(legend.position = "bottom")
```

# Dosing and INR Response

```{r}
daily <- data_daily %>%
    ungroup() %>%
    filter(
        warfarin_day <= 7,
        med.datetime >= start_date,
        med.datetime <= end_date
    ) %>%
    left_join(
        data_warfarin[c("millennium.id", "group", "initiation", "indication_group")], 
        by = "millennium.id"
    ) %>%
    mutate_at("group", factor, levels = c("traditional", "pharmacy")) %>%
    mutate_at("warfarin_day", as.integer) %>%
    mutate_at("warfarin_day", factor) 

pharm <- daily %>%
    filter(group == "pharmacy") %>%
    distinct(millennium.id) %>%
    nrow()

trad <- daily %>%
    filter(group == "traditional") %>%
    distinct(millennium.id) %>%
    nrow()
```

There were `r format(pharm, big.mark = ",")` patients in the Pharmacy Dosing Service group, and `r format(trad, big.mark = ",")` patients in the traditional management group. In comparing the warfarin doses administered on each of the first 7 days of therapy, there was a higher initial dose ordered by the Pharmacy Dosing Service (see Figure 4).

```{r, fig.cap="Distribution of warfarin dose by day of therapy. The p-value is < 0.01 for each day of warfarin."}
p <- daily %>%
    group_by(warfarin_day) %>%
    do(tidy(wilcox.test(med.dose ~ group, data = .))) %>%
    mutate_at("p.value", round, digits = 3)

ggplot(daily, aes(x = warfarin_day, y = med.dose, fill = group)) +
    geom_boxplot() +
    xlab("Day of therapy") +
    scale_y_continuous("Warfarin dose (mg)",  breaks = seq(0, 30, 2.5)) +
    # scale_fill_brewer(NULL, palette = "Greys", labels = c("Traditional", "Pharmacy")) +
    my_scale_fill() +
    theme_bg(base_family = "serif") +
    theme(legend.position = "bottom")
```

Patients managed by the Pharmacy Dosing Service had a greater rise in INR over the first 7 days of therapy, and were more likely to reach an INR around 2 by day 5. 

```{r, fig.cap="INR response during the first 7 days of warfarin therapy. The p-value was < 0.001 for all days except day 3 (p = 0.002) and day 4 (p = 0.869). Patients with an INR target below 2 were excluded from this figure. The usual therapeutic INR range of 2 to 3 is shaded for reference."}
inr <- data_daily %>%
    left_join(data_warfarin[c("millennium.id", "goal.low")], by = "millennium.id") %>%
    ungroup() %>%
    filter(
        warfarin_day <= 7,
        med.datetime >= start_date,
        med.datetime <= end_date,
        !is.na(inr),
        goal.low >= 2
    ) %>%
    left_join(
        data_warfarin[c("millennium.id", "group", "initiation", "indication_group")], 
        by = "millennium.id") %>%
    mutate_at("warfarin_day", as.numeric) 

p <- inr %>%
    mutate_at("warfarin_day", round, digits = 0) %>%
    group_by(warfarin_day) %>%
    do(tidy(wilcox.test(inr ~ group, data = .))) %>%
    mutate_at("p.value", round, digits = 3)

ggplot(inr, aes(x = warfarin_day, y = inr, linetype = group)) +
    geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = 2, ymax = 3), fill = "gray95") +
    geom_smooth(color = "black") +
    scale_x_continuous("Day of therapy", breaks = seq(1, 7, 1)) +
    ylab("INR") +
    scale_linetype_discrete(NULL, labels = c("Pharmacy", "Traditional")) +
    coord_cartesian(ylim = c(1, 3.5)) +
    theme_bg(base_family = "serif") +
    theme(legend.position = "bottom")
```

There was a greater time-in-therapeutic range for patients managed by the Pharmacy Dosing Service in FY17 compared with traditional management. In a historical comparison, there was no difference in the time-in-therapeutic range for patients managed by the Pharmacy Dosing Service in FY17 vs. FY16.

```{r, fig.cap="Percent time in therapeutic range among patients managed by the Pharmacy Dosing Service in FY17 vs. Traditional management, as well as patients managed by the Pharmacy Dosing Service in FY17 vs. FY16. The p-value for each comparison is included on the figure.", cache=TRUE}
p1 <- t.test(ttr_pharm_curr$perc.time, ttr_trad_curr$perc.time, var.equal = TRUE)
# p2 <- t.test(ttr_pharm17$perc.time, ttr_pharm16$perc.time, var.equal = TRUE)

ttr_pharm_curr %>%
    bind_rows(ttr_trad_curr) %>%
    # bind_rows(ttr_trad17, ttr_pharm16) %>%
    mutate_at("group", factor, levels = c("pharmacy", "traditional")) %>%
    ggplot(aes(x = group, y = perc.time)) +
    geom_segment(aes(x = 1.1, xend = 1.9, y = 50, yend = 50), color = "grey35", size = 0.5) +
    # geom_segment(aes(x = 1.1, xend = 2.9, y = 75, yend = 75), color = "grey35", size = 0.5) +
    geom_label(aes(x = 1.5, y = 50, label = round(p1$p.value, 2)), size = 3, color = "grey35") +
    # geom_label(aes(x = 2.5, y = 75, label = round(p2$p.value, 2)), size = 3, color = "grey35") +
    geom_boxplot() +
    scale_x_discrete(NULL, labels = c("Pharmacy", "Traditional")) +
    ylab("Time in Therapeutic Range (%)") +
    theme_bg(xticks = FALSE, base_family = "serif")
```


```{r, eval=FALSE}
t.test(ttr_pharm17$perc.time, ttr_pharm16$perc.time, var.equal = TRUE) %>%
    tidy() %>%
    rename(FY17 = estimate1, FY16 = estimate2) %>%
    knitr::kable(digits = 3, caption = "Percent time in therapeutic range, FY17 pharmacy vs. FY16 pharmacy")
```

# Outcomes and Events

There were no differences in bleeding, need for anticoagulation reversal, readmission within 30 dyas, or other adverse events (see Table 1).`r margin_note("Unplanned return included any ER visit, observation, or inpatient stay at any MH hospital within 30 days of discharge from MH-TMC.")` 

`r margin_note("Hemoglobin decrease of >/= 2 g/dL within a 48-hour period")`
`r margin_note("Reversal medications included phytonadione, 4-factor PCC, factor VIIa, aminocaproic acid, or tranexamic acid")`

```{r}
hgb <- data_hgb_drop %>%
    distinct(millennium.id, .keep_all = TRUE) %>%
    full_join(
        data_warfarin[c("millennium.id", "group", "warfarin_start")], 
        by = "millennium.id"
    ) %>%
    filter(warfarin_start >= start_date,
           warfarin_start <= end_date) 
    
tbl_hgb <- hgb %>%
    mutate(hgb = !is.na(hgb_day)) %>%
    group_by(group) %>%
    summarize_at("hgb", mean, na.rm = TRUE) %>%
    mutate_at("hgb", funs(. * 100)) %>%
    spread(group, hgb) %>%
    mutate(Event = "Hemoglobin decrease, 48-hours") 

tmp_hgb <- hgb %>%
    mutate(hgb = if_else(!is.na(hgb_day), "hgb", "no_hgb")) %>%
    mutate_at(c("group", "hgb"), factor)

p_hgb <- chisq.test(tmp_hgb$group, tmp_hgb$hgb)
```

```{r}
prbc <- data_prbc %>%
    distinct(millennium.id, .keep_all = TRUE) %>%
    full_join(
        data_warfarin[c("millennium.id", "group", "warfarin_start")], 
        by = "millennium.id"
    ) %>%
    filter(
        warfarin_start >= start_date,
        warfarin_start <= end_date
    ) 

tbl_prbc <- prbc %>%
    mutate(prbc = !is.na(prbc_day)) %>%
    group_by(group) %>%
    summarize_at("prbc", mean, na.rm = TRUE) %>%
    mutate_at("prbc", funs(. * 100)) %>%
    spread(group, prbc) %>%
    mutate(Event = "Transfusion PRBC") 

tmp_prbc <- prbc %>%
    mutate(prbc = if_else(!is.na(prbc_day), "prbc", "no_prbc")) %>%
    mutate_at(c("group", "prbc"), factor)

p_prbc <- chisq.test(tmp_prbc$group, tmp_prbc$prbc)
```

```{r}
ffp <- data_ffp %>%
    distinct(millennium.id, .keep_all = TRUE) %>%
    full_join(
        data_warfarin[c("millennium.id", "group", "warfarin_start")], 
        by = "millennium.id"
    ) %>%
    filter(
        warfarin_start >= start_date,
        warfarin_start <= end_date
    ) 

tbl_ffp <- ffp %>%
    mutate(ffp = !is.na(ffp_day)) %>%
    group_by(group) %>%
    summarize_at("ffp", mean, na.rm = TRUE) %>%
    mutate_at("ffp", funs(. * 100)) %>%
    spread(group, ffp) %>%
    mutate(Event = "Transfusion FFP") 

tmp_ffp <- ffp %>%
    mutate(ffp = if_else(!is.na(ffp_day), "ffp", "no_ffp")) %>%
    mutate_at(c("group", "ffp"), factor)

p_ffp <- chisq.test(tmp_ffp$group, tmp_ffp$ffp)
```

```{r}
reversal <- data_reversal %>%
    distinct(millennium.id, .keep_all = TRUE) %>%
    full_join(
        data_warfarin[c("millennium.id", "group", "warfarin_start")],
        by = "millennium.id"
    ) %>%
    filter(
        warfarin_start >= start_date,
        warfarin_start <= end_date
    ) 

tbl_reversal <- reversal %>%
    mutate(reversal = !is.na(rev_day)) %>%
    group_by(group) %>%
    summarize_at("reversal", mean, na.rm = TRUE) %>%
    mutate_at("reversal", funs(. * 100)) %>%
    spread(group, reversal) %>%
    mutate(Event = "Reversal Medication") 

tmp_reversal <- reversal %>%
    mutate(reversal = if_else(!is.na(rev_day), "reversal", "no_reversal")) %>%
    mutate_at(c("group", "reversal"), factor)

p_reversal <- chisq.test(tmp_reversal$group, tmp_reversal$reversal)
```

```{r}
procedures <- data_procedures %>%
    distinct(millennium.id, .keep_all = TRUE) %>%
    full_join(
        data_warfarin[c("millennium.id", "group", "warfarin_start")],
        by = "millennium.id"
    ) %>%
    filter(
        warfarin_start >= start_date,
        warfarin_start <= end_date
    ) 

tbl_procedures <- procedures %>%
    mutate(procedures = !is.na(warf_day)) %>%
    group_by(group) %>%
    summarize_at("procedures", mean, na.rm = TRUE) %>%
    mutate_at("procedures", funs(. * 100)) %>%
    spread(group, procedures) %>%
    mutate(Event = "Procedure") 

tmp_procedures <- procedures %>%
    mutate(procedures = if_else(!is.na(warf_day), "procedure", "no_procedure")) %>%
    mutate_at(c("group", "procedures"), factor)

p_procedures <- chisq.test(tmp_procedures$group, tmp_procedures$procedures)
```

```{r}
revisit_all <- data_warfarin %>%
    select(millennium.id, group, warfarin_start) %>%
    left_join(
        data_revisits[c(
            "millennium.id", 
            "next_encounter", 
            "revisit_facility", 
            "revisit_visit_type"
        )], 
        by = "millennium.id"
    ) %>%
    filter(
        warfarin_start >= start_date,
        warfarin_start <= end_date
    ) 

tbl_revisit <- revisit_all %>%
    mutate(revisit = !is.na(revisit_facility)) %>%
    group_by(group) %>%
    summarize_at("revisit", mean, na.rm = TRUE) %>%
    mutate_at("revisit", funs(. * 100)) %>%
    spread(group, revisit) %>%
    mutate(Event = "Unplanned return, 30 days") 

tmp_revisit <- revisit_all %>%
    mutate(revisit = if_else(!is.na(revisit_facility), "revisit", "no_revisit")) %>%
    mutate_at(c("group", "revisit"), factor)

p_revisit <- chisq.test(tmp_revisit$group, tmp_revisit$revisit)
```

```{r}
tbl_readmit <- revisit_all %>%
    mutate(readmit = revisit_visit_type == "Inpatient" & !is.na(revisit_facility)) %>%
    group_by(group) %>%
    summarize_at("readmit", mean, na.rm = TRUE) %>%
    mutate_at("readmit", funs(. * 100)) %>%
    spread(group, readmit) %>%
    mutate(Event = "Unplanned readmission, 30 days") 

tmp_readmit <- revisit_all %>%
    mutate(
        readmit = if_else(
            revisit_visit_type == "Inpatient", 
            "readmit", 
            "no_readmit", 
            "no_readmit"
        )
    ) %>%
    mutate_at(c("group", "readmit"), factor)

p_readmit <- chisq.test(tmp_readmit$group, tmp_readmit$readmit)
```

```{r}
p <- c(
    p_hgb$p.value, 
    p_prbc$p.value,
    p_ffp$p.value, 
    p_reversal$p.value,
    p_procedures$p.value, 
    p_revisit$p.value, 
    p_readmit$p.value
)

tbl_hgb %>%
    bind_rows(
        tbl_prbc, 
        tbl_ffp, 
        tbl_reversal, 
        tbl_procedures, 
        tbl_revisit, 
        tbl_readmit
    ) %>%
    select(Event, Pharmacy = pharmacy, Traditional = traditional) %>%
    mutate(`p-value` = p) %>%
    knitr::kable(
        digits = c(0, 1, 1, 2), 
        caption = "Outcomes and Events During Warfarin Therapy"
    )
```

# Conclusion

The Pharmacy Dosing Service continued to provide safe and effective management of warfarin for patients in fiscal year 2018. 
