---
title: "Pharmacy-managed warfarin dosing service analysis for fiscal year 2019"
author: "Brian Gulbis, PharmD, BCPS"
date: "October 28, 2019"
output:
  tufte::tufte_handout:
    latex_engine: xelatex
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(
    tidy = FALSE, 
    echo = FALSE, 
    message = FALSE, 
    warning = FALSE, 
    cache.extra = packageVersion('tufte'), 
    kable.force.latex = TRUE
)

options(htmltools.dir.version = FALSE)
```

```{r data, message=FALSE}
library(data.table)
library(tidyverse)
library(lubridate)
library(broom)
library(mbohelpr)
library(themebg)
library(kableExtra)

data_dir <- "../data/tidy/fy2019"

find_ind <- function(x, reg) {
    stringr::str_detect(x, regex(reg, ignore_case = TRUE))
}


data_consult_orders <- get_data(data_dir, "consult_orders") %>%
    filter(str_detect(nurse_unit, "^HH|^HVI")) %>%
    mutate(
        order_month = floor_date(order_datetime, unit = "month"),
        fiscal_year = year(order_month %m+% months(6))
    )
    
data_consult_tasks <- get_data(data_dir, "consult_tasks") %>%
    filter(
        str_detect(nurse_unit, "^HH|^HVI"),
        task_date < ymd("2019-07-01")
    ) %>%
    mutate(
        task_month = floor_date(task_date, unit = "month"),
        fiscal_year = year(task_month %m+% months(6))
    )

data_demographics <- get_data(data_dir, "demographics")

data_doac_doses <- get_data(data_dir, "doac_doses") %>%
    filter(str_detect(nurse_unit, "^HH|^HVI")) %>%
    mutate(
        med_month = floor_date(med_datetime, unit = "month"),
        fiscal_year = year(med_month %m+% months(6))
    )
    
data_labs <- get_data(data_dir, "labs")
data_measures <- get_data(data_dir, "measures")
data_reencounters <- get_data(data_dir, "reencounters")
data_reversal_meds <- get_data(data_dir, "reversal_meds")
data_warfarin_details <- get_data(data_dir, "warfarin_details")

data_warfarin_doses <- get_data(data_dir, "warfarin_doses") %>%
    filter(str_detect(nurse_unit, "^HH|^HVI")) %>%
    mutate(
        med_day = floor_date(med_datetime, unit = "day"),
        med_month = floor_date(med_datetime, unit = "month"),
        fiscal_year = year(med_month %m+% months(6))
    )

data_warfarin_home_meds <- get_data(data_dir, "warfarin_home_meds")
data_warfarin_orders <- get_data(data_dir, "warfarin_orders")

df_indications <- data_warfarin_details %>%
    filter(detail == "Warfarin Indication") %>%
    mutate(
        indication = case_when(
            find_ind(
                result_value, 
                "Atrial fibrillation|a(.*)?fib|a(.*)?flutter"
            ) ~ "afib",
            find_ind(
                result_value, 
                "Deep vein thrombosis|DVT(?!( prophylaxis))|VTE"
            ) ~ "dvt",
            find_ind(
                result_value, 
                "Pulmonary embolism|PE"
            ) ~ "pe",
            find_ind(
                result_value, 
                "Heart valve \\(Mech/porc/bioprost\\)|valve|avr|mvr"
            ) ~ "valve",
            find_ind(
                result_value, 
                "st(ro|or)ke|cva|ica|mca"
            ) ~ "stroke",
            find_ind(
                result_value, 
                "vad|hm[ ]?ii|heart( )?mate|heartware|syncardia|total artificial heart|tah"
            ) ~ "vad",
            find_ind(
                result_value, 
                "throm|clot|emboli|occl"
            ) ~ "thrombus",
            find_ind(
                result_value, 
                "malig|anti(.)?phos|lupus|apla|hypercoag|deficien|leiden|fvl|factor v"
            ) ~ "hypercoag",
            find_ind(
                result_value, 
                "prophylax"
            ) ~ "prophylax",
            TRUE ~ "other"
        )
    ) %>%
    distinct(encounter_id, indication) 
    # mutate(value = TRUE) %>%
    # spread(indication, value, fill = FALSE)
        
df_inr_goal <- data_warfarin_details %>%
    filter(
        detail == "INR Range",
        result_value != ""
    ) %>%
    mutate(inr_range = result_value) %>%
    mutate_at(
        "inr_range", 
        str_replace_all, 
        pattern = " |", 
        replacement = ""
    ) %>%
    mutate_at(
        "inr_range", 
        str_replace_all, 
        pattern = regex("to|--|/|=|,|-\\.", ignore_case = TRUE), 
        replacement = "-"
    ) %>%
     mutate_at(
        "inr_range",
        str_replace_all,
        pattern = ">1.5-<2.5",
        replacement = "1.5-2.5"
    ) %>%
    mutate_at(
        "inr_range",
        str_replace_all,
        pattern = regex(
            c(
                # "^1.[5-9]$" = "1.5-2",
                "^2$|^2.0$" = "1.5-2.5",
                "^2\\.[1-4]$" = "2-2.5",
                "^2.5$" = "2-3",
                # "^2.[6-9]$" = "2.5-3",
                "^3$" = "2.5-3.5",
                # "^3.5$" = "3-4",
                # "^22" = "2",
                # "2.53.5" = "2.5-3.5",
                "^23$" = "2-3",
                "30$" = "3",
                "^1-5" = "1.5"
            ),
            ignore_case = TRUE
        )
    ) %>%
    extract(
        inr_range,
        c("goal_low", "goal_high"), 
        regex = "([0-9\\.]+ ?)-( ?[0-9\\.]+)",
        remove = FALSE, 
        convert = TRUE
    ) %>%
    filter(
        !is.na(goal_low),
        goal_low < goal_high
    ) %>%
    mutate_at(c("goal_low", "goal_high"), as.numeric) %>%
    select(-result_value, -inr_range)

```

```{r}
df_consult_pts <- data_consult_orders %>%
    arrange(encounter_id, order_datetime) %>%
    distinct(encounter_id, .keep_all = TRUE) %>%
    mutate(
        consult = TRUE,
        consult_date = floor_date(order_datetime, unit = "day")
    ) %>%
    select(
        encounter_id, 
        consult_date,
        nurse_unit_consult = nurse_unit,
        med_service, 
        consult
    )

df_warf_start <- data_warfarin_doses %>%
    arrange(encounter_id, med_datetime) %>%
    distinct(encounter_id, .keep_all = TRUE) %>%
    mutate(warfarin_start = floor_date(med_datetime, unit = "day")) %>%
    select(
        encounter_id,
        med_month,
        warfarin_start, 
        nurse_unit_warfarin = nurse_unit, 
        fiscal_year
    )

df_demog <- data_demographics %>%
    inner_join(df_warf_start, by = "encounter_id") %>%
    left_join(df_consult_pts, by = "encounter_id") %>%
    mutate_at("consult", list(~coalesce(., FALSE))) %>%
    mutate(
        consult_day = difftime(
            consult_date, 
            warfarin_start, 
            units = "days"
        )
    )

df_warf_month <- data_warfarin_doses %>%
    count(med_month, name = "warfarin") %>%
    mutate_at("med_month", as_date)

df_consult_month <- data_consult_tasks %>%
     count(task_month, name = "consults")

df_doac_month <- data_doac_doses %>%
    mutate(med_day = floor_date(med_datetime, unit="day")) %>%
    distinct(encounter_id, medication, med_month, med_day) %>%
    count(med_month, name = "doac") %>%
    mutate_at("med_month", as_date)
```

```{r}
df_warf_ord <- data_warfarin_orders %>%
    rename(
        nurse_unit_order = nurse_unit,
        med_service_order = med_service
    ) %>%
    arrange(encounter_id, order_datetime) %>%
    mutate(order_date = floor_date(order_datetime, unit = "day")) %>%
    inner_join(df_warf_start, by = c("encounter_id", "order_date" = "warfarin_start")) %>%
    left_join(df_consult_pts, by = "encounter_id") %>%
    mutate_at("consult", list(~coalesce(., FALSE)))

df_warf_duration <- data_warfarin_doses %>%
    arrange(encounter_id, med_datetime) %>%
    group_by(encounter_id) %>%
    summarize_at("med_day", list(warfarin_start = first, warfarin_stop = last))

 df_doses_daily <- data_warfarin_doses %>%
    arrange(encounter_id, med_datetime) %>%
    group_by(encounter_id, fiscal_year, med_day) %>%
    summarize_at("dose", sum, na.rm = TRUE) %>%
    group_by(encounter_id, fiscal_year) %>%
    mutate(warf_day = difftime(med_day, first(med_day), units = "days") + 1) %>%
    mutate_at("warf_day", as.integer) %>%
    filter(warf_day <= 10) %>%
    ungroup() %>%
    select(encounter_id, fiscal_year, warf_day, dose) 

df_inr_daily <- df_warf_duration %>%
    inner_join(data_labs, by = "encounter_id") %>%
    filter(lab == "INR") %>%
    mutate(
        lab_day = floor_date(lab_datetime, unit = "days"),
        lab_month = floor_date(lab_datetime, unit = "month"),
        warf_day = difftime(lab_day, warfarin_start, units = "days") + 1,
        fiscal_year = year(lab_month %m+% months(6))
    ) %>%
    mutate_at("warf_day", as.integer) %>%
    filter(
        lab_day >= warfarin_start,
        lab_day <= warfarin_stop,
        warf_day <= 10
    ) %>%
    group_by(encounter_id, fiscal_year, lab_day, warf_day) %>%
    summarize_at("result_value", max, na.rm = TRUE) %>%
    ungroup() %>%
    select(encounter_id, fiscal_year, warf_day, inr = result_value) 

df_consult_daily <- df_warf_duration %>%
    mutate_at(c("warfarin_start", "warfarin_stop"), as_date) %>%
    inner_join(data_consult_tasks, by = "encounter_id") %>%
    mutate(
        warf_day = difftime(task_date, warfarin_start, units = "days") + 1,
        task_month = floor_date(task_date, unit = "month"),
        fiscal_year = year(task_month %m+% months(6))
    ) %>%
    mutate_at("warf_day", as.integer) %>%
    filter(
        task_date >= warfarin_start,
        task_date <= warfarin_stop,
        warf_day <= 10
    ) %>%
    select(encounter_id, fiscal_year, warf_day) %>%
    mutate(consult = TRUE)

df_doses_inr <- df_doses_daily %>%
    full_join(
        df_inr_daily, 
        by = c("encounter_id", "fiscal_year", "warf_day")
    ) %>%
    full_join(
        df_consult_daily,
        by = c("encounter_id", "fiscal_year", "warf_day")
    ) %>%
    arrange(encounter_id, warf_day)

df_goal_eq2_3 <- df_inr_goal %>%
    filter(
        goal_low == 2,
        goal_high == 3
    )

```

# Summary

Utilization of the Pharmacy-Managed Warfarin Dosing Service continued to increase in FY18. The areas with the highest utilization were Rehab and some areas of HVI. Patients managed by the pharmacy service had a small but statistically significantly greater time in therapeutic range than patients with traditional warfarin management.

# Service Utilization

There has been a % increase in the number of warfarin patients being managed by the Pharmacy Dosing Service over the past two fiscal years (see Figure 1). Conversely, there has been a % decrease in the number of patients with traditional warfarin management (warfarin managed by a physician or advanced practice provider) during this time. 

```{r, fig.cap="The number of monthly doses of warfarin, number of monthly warfarin doses managed by the Pharmacy Dosing Service, and the number of monthly doses of direct oral anticoagulants (DOACs) over the past three fiscal years."}

df_warf_month %>%
    inner_join(df_consult_month, by = c("med_month" = "task_month")) %>%
    inner_join(df_doac_month, by = "med_month") %>%
    gather(key, value, -med_month) %>%
    ggplot(aes(x = med_month, y = value, color = key, size = key)) +
    annotate("text", x = mdy("9/1/2016"), y = 950, label = "All warfarin\npatients", family = "serif", color = "grey25") +
    annotate("text", x = mdy("9/1/2016"), y = 750, label = "Pharmacy\nconsults", family = "serif", color = "black") +
    annotate("text", x = mdy("9/1/2016"), y = 250, label = "DOACs", family = "serif", color = "grey50") +
    geom_line() +
    ggtitle("Monthly doses of oral anticoagulants") +
    scale_x_date(NULL, date_breaks = "4 months", date_labels = "%b %y") +
    scale_y_continuous("Doses") +
    scale_color_manual(NULL, values = c("black", "grey50", "grey25")) +
    scale_size_manual(NULL, values = c(1, 0.5, 0.5)) +
    theme_bg_print(base_family = "serif") +
    theme(legend.position = "none")

```

The highest volume of warfarin use occured in the Heart Failure IMU, followed by CIMU and CVIMU (see Figure 2). Utilization of the Pharmacy Dosing Service remains low in HFIMU, although Clinical Pharmacists and Clinical Pharmacist-Specialists participate in the management of warfarin in this area.



```{r, fig.cap="The number of orders for warfarin by the primary service at the time of warfarin initiation.", fig.height=3.5}
df_warf_ord %>%
    filter(fiscal_year == 2019) %>%
    mutate_at(
        "med_service_order",
        str_replace_all,
        pattern = c(
            "Thoracic/Cardiac Sur Service" = "CV Surgery",
            "Therapy - Phys/Occ/Speech/Cardiac" = "Rehab",
            "Pulmonology/Respiratory Therapy" = "Pulmonary"
        )
    ) %>%
    mutate_at("med_service_order", fct_infreq) %>%
    mutate_at("med_service_order", fct_lump, n = 10) %>%
    mutate_at("med_service_order", fct_rev) %>%
    ggplot(aes(x = med_service_order, fill = consult)) +
    geom_bar(position = "dodge") +
    ggtitle("Warfarin utilization by primary service") +
    xlab(NULL) +
    ylab("Number of warfarin patients") +
    scale_fill_manual(
        NULL, 
        values = c("grey50", "black"), 
        labels = c("Traditional", "Pharmacy")
    ) +
    coord_flip() +
    theme_bg_print(base_family = "serif") +
    theme(legend.position = "top")
```


The Pharmacy Dosing Service managed a higher percent of patients with venous thromboembolism and mechanical heart valves, while the traditional management was responsible for a higher percent of patients with mechanical circulatory support devices (see Figure 3). There was a similar percent of patients with atrial fibrillation in both groups.

```{r, fig.cap="Warfarin indications", fig.height=3.5}
df_demog %>%
    filter(fiscal_year == 2019) %>%
    inner_join(df_indications, by = "encounter_id") %>%
    mutate_at(
        "indication",
        str_replace_all,
        pattern = c(
            "afib" = "A.fib",
            "vad" = "VAD",
            "valve" = "Valve",
            "dvt" = "DVT",
            "hypercoag" = "Hypercoaguable",
            "pe" = "PE",
            "thrombus" = "Thrombus",
            "stroke" = "Stroke",
            "other" = "Other"
        )
    ) %>% 
    mutate_at("indication", fct_infreq) %>%
    mutate_at("indication", fct_rev) %>%
    mutate_at("fiscal_year", as_factor) %>%
    ggplot(aes(x = indication, fill = consult)) +
    geom_bar(position = "dodge") +
    ggtitle("Warfarin indications") +
    xlab(NULL) +
    ylab("Patients (%)") + # calculate percent?
    scale_fill_manual(
        NULL, 
        values = c("grey50", "black"), 
        labels = c("Traditional", "Pharmacy")
    ) +
    coord_flip() +
    theme_bg_print(base_family = "serif") +
    theme(legend.position = "top")
```

# Dosing and INR Response

```{r}
# daily <- data_daily %>%
#     ungroup() %>%
#     filter(
#         warfarin_day <= 7,
#         med.datetime >= start_date,
#         med.datetime <= end_date
#     ) %>%
#     left_join(
#         data_warfarin[c("millennium.id", "group", "initiation", "indication_group")], 
#         by = "millennium.id"
#     ) %>%
#     mutate_at("group", factor, levels = c("traditional", "pharmacy")) %>%
#     mutate_at("warfarin_day", as.integer) %>%
#     mutate_at("warfarin_day", factor) 
# 
# pharm <- daily %>%
#     filter(group == "pharmacy") %>%
#     distinct(millennium.id) %>%
#     nrow()
# 
# trad <- daily %>%
#     filter(group == "traditional") %>%
#     distinct(millennium.id) %>%
#     nrow()

# `r format(pharm, big.mark = ",")`
```

There were XX patients in the Pharmacy Dosing Service group, and patients in the traditional management group. In comparing the warfarin doses administered on each of the first 7 days of therapy, there was a higher median daily dose ordered for patients managed by the Pharmacy Dosing Service (see Figure 4).

```{r, fig.cap="Distribution of warfarin dose by day of therapy. The thick horizontal black line represents the median dose for each group each day. The p-value is < 0.05 for each day of warfarin except day 2 where there was no significant difference."}
p <- df_doses_inr %>%
    semi_join(df_goal_eq2_3, by = "encounter_id") %>%
    filter(
        warf_day <= 7,
        fiscal_year == 2019
    ) %>%
    mutate_at("consult", list(~coalesce(., FALSE))) %>%
    group_by(warf_day) %>%
    do(tidy(wilcox.test(dose ~ consult, data = .))) %>%
    mutate_at("p.value", round, digits = 3)


df_doses_inr %>%
    semi_join(df_goal_eq2_3, by = "encounter_id") %>%
    filter(
        warf_day <= 7,
        fiscal_year == 2019
    ) %>%
    mutate_at("warf_day", as_factor) %>%
    mutate_at("consult", list(~coalesce(., FALSE))) %>%
    ggplot(aes(x = warf_day, y = dose, fill = consult, color = consult)) +
    geom_boxplot(outlier.shape = NA) +
    ggtitle("Warfarin daily doses") +
    xlab("Day of therapy") +
    ylab("Dose (mg)") +
    scale_color_manual(
        NULL,
        values = c("grey25", "grey50"),
        labels = c("Traditional", "Pharmacy")
    ) + 
    scale_fill_manual(
        NULL, 
        values = c("grey50", "black"), 
        labels = c("Traditional", "Pharmacy")
    ) +
    coord_cartesian(ylim = c(0, 15)) +
    theme_bg_print(base_family = "serif") +
    theme(legend.position = "top")

```

Patients managed by the Pharmacy Dosing Service had a greater rise in INR over the first 7 days of therapy, and were more likely to reach an INR around 2 by day 5 (see Figure 5). 

```{r, fig.cap="There was a larger increase in the INR over the first 10 days of warfarin therapy in patients managed by the Pharmacy Dosing Service compared with traditional provider management. The p-value was < 0.05 on all days except days 4 and 5 where there was no significant difference. Only patients with an INR target of 2 to 3 were included."}

p <- df_doses_inr %>%
    semi_join(df_goal_eq2_3, by = "encounter_id") %>%
    filter(
        # warf_day <= 7
        fiscal_year == 2019,
        !is.na(inr)
    ) %>%
    mutate_at("consult", list(~coalesce(., FALSE))) %>%
    mutate_at("warf_day", round, digits = 0) %>%
    group_by(warf_day) %>%
    do(tidy(wilcox.test(inr ~ consult, data = .))) %>%
    mutate_at("p.value", round, digits = 3)

df_doses_inr %>%
    semi_join(df_goal_eq2_3, by = "encounter_id") %>%
    filter(
        # warf_day <= 7
        fiscal_year == 2019,
        !is.na(inr)
    ) %>%
    mutate_at("consult", list(~coalesce(., FALSE))) %>%
    ggplot(aes(x = warf_day, y = inr, color = consult)) +
    geom_hline(yintercept = 2, linetype = "dashed", color = "grey75") +
    annotate("text", x = 1.5, y = 1.4, label = "Pharmacy", color = "black", family = "serif") + # adjust font face and size
    annotate("text", x = 1.5, y = 1.8, label = "Traditional", color = "grey50", family = "serif") +
    geom_smooth() +
    ggtitle("INR response") +
    scale_x_continuous("Day of therapy", breaks = seq(1, 10, 1)) +
    ylab("INR") +
    scale_color_manual(
        NULL,
        values = c("grey50", "black"),
        labels = c("Traditional", "Pharmacy")
    ) + 
    coord_cartesian(ylim = c(1, 3)) +
    theme_bg_print(base_family = "serif") +
    theme(legend.position = "none")

```

# Outcomes and Events

There were no differences in bleeding, need for anticoagulation reversal, readmission within 30 dyas, or other adverse events between groups (see Table 1).

```{r high_inr}
df_high_inr <- df_doses_inr %>%
    filter(
        inr >= 5
    ) %>%
    distinct(encounter_id) %>%
    mutate(high_inr = TRUE)
```

```{r hgb_drop}
df_hgb <- df_warf_duration %>%
    inner_join(data_labs, by = "encounter_id") %>%
    arrange(encounter_id, lab_datetime) %>%
    filter(lab == "Hgb") %>%
    mutate(
        lab_day = floor_date(lab_datetime, unit = "days"),
        lab_month = floor_date(lab_datetime, unit = "month"),
        warf_day = difftime(lab_day, warfarin_start, units = "days") + 1,
        fiscal_year = year(lab_month %m+% months(6))
    ) %>%
    mutate_at("warf_day", as.integer) %>%
    filter(
        lab_day >= warfarin_start,
        lab_day <= warfarin_stop + 1,
    ) 

dt_hgb <- as.data.table(df_hgb)

dt_window <- dt_hgb[, .(
    encounter_id, 
    window_low = lab_datetime - hours(48), 
    window_high = lab_datetime
)]

dt_hgb_max <- dt_hgb[
    dt_window, 
    on=.(
        encounter_id, 
        lab_datetime >= window_low, 
        lab_datetime <= window_high
    ), 
    .(max_hgb = max(result_value)), 
    by=.EACHI
    ][, lab_datetime := NULL
      ][, head(.SD, 1), by = c("encounter_id", "lab_datetime")]

dt_hgb_drop <- dt_hgb[dt_hgb_max, on=.(encounter_id, lab_datetime)]

df_hgb_drop <- as_tibble(dt_hgb_drop) %>%
    arrange(encounter_id, lab_datetime) %>%
    mutate(hgb_chg = result_value - max_hgb) %>%
    filter(
        hgb_chg <= -2,
        warf_day <= 10
    ) %>% 
    distinct(encounter_id) %>%
    mutate(hgb_drop = TRUE)

```

```{r reversal}
df_reversal <- data_reversal_meds %>%
    inner_join(df_warf_duration, by = "encounter_id") %>%
    filter(
        med_datetime >= warfarin_start,
        med_datetime <= warfarin_stop 
    ) %>%
    distinct(encounter_id) %>%
    mutate(reversal_med = TRUE)
```

```{r revisits}
df_revisit <- df_demog %>%
    left_join(data_reencounters, by = "encounter_id") %>%
    distinct(encounter_id) %>%
    mutate(revisit = TRUE)
```

```{r readmits}
df_readmit <- df_demog %>%
    left_join(data_reencounters, by = "encounter_id") %>%
    filter(readmit_type == "Inpatient") %>%
    distinct(encounter_id) %>%
    mutate(readmit = TRUE)
```

```{r}
df_outcomes <- df_demog %>%
    left_join(df_high_inr, by = "encounter_id") %>%
    left_join(df_hgb_drop, by = "encounter_id") %>%
    left_join(df_reversal, by = "encounter_id") %>%
    left_join(df_revisit, by = "encounter_id") %>%
    left_join(df_readmit, by = "encounter_id") %>%
    filter(fiscal_year == 2019) %>%
    select(encounter_id, consult, disch_disposition, high_inr:readmit) %>%
    mutate_at(
        c("high_inr", "hgb_drop", "reversal_med", "revisit", "readmit"),
        list(~coalesce(., FALSE))
    )
```


```{r, eval=FALSE}
prbc <- data_prbc %>%
    distinct(millennium.id, .keep_all = TRUE) %>%
    full_join(
        data_warfarin[c("millennium.id", "group", "warfarin_start")], 
        by = "millennium.id"
    ) %>%
    filter(
        warfarin_start >= start_date,
        warfarin_start <= end_date
    ) 

tbl_prbc <- prbc %>%
    mutate(prbc = !is.na(prbc_day)) %>%
    group_by(group) %>%
    summarize_at("prbc", mean, na.rm = TRUE) %>%
    mutate_at("prbc", funs(. * 100)) %>%
    spread(group, prbc) %>%
    mutate(Event = "Transfusion PRBC") 

tmp_prbc <- prbc %>%
    mutate(prbc = if_else(!is.na(prbc_day), "prbc", "no_prbc")) %>%
    mutate_at(c("group", "prbc"), factor)

p_prbc <- chisq.test(tmp_prbc$group, tmp_prbc$prbc)
```

```{r, eval=FALSE}
ffp <- data_ffp %>%
    distinct(millennium.id, .keep_all = TRUE) %>%
    full_join(
        data_warfarin[c("millennium.id", "group", "warfarin_start")], 
        by = "millennium.id"
    ) %>%
    filter(
        warfarin_start >= start_date,
        warfarin_start <= end_date
    ) 

tbl_ffp <- ffp %>%
    mutate(ffp = !is.na(ffp_day)) %>%
    group_by(group) %>%
    summarize_at("ffp", mean, na.rm = TRUE) %>%
    mutate_at("ffp", funs(. * 100)) %>%
    spread(group, ffp) %>%
    mutate(Event = "Transfusion FFP") 

tmp_ffp <- ffp %>%
    mutate(ffp = if_else(!is.na(ffp_day), "ffp", "no_ffp")) %>%
    mutate_at(c("group", "ffp"), factor)

p_ffp <- chisq.test(tmp_ffp$group, tmp_ffp$ffp)
```


```{r, eval=FALSE}
p <- c(
    p_hgb$p.value, 
    p_vizient$p.value,
    p_prbc$p.value,
    p_ffp$p.value, 
    p_reversal$p.value,
    p_procedures$p.value, 
    p_revisit$p.value, 
    p_readmit$p.value
)

tbl_hgb %>%
    bind_rows(
        tbl_vizient,
        tbl_prbc, 
        tbl_ffp, 
        tbl_reversal, 
        tbl_procedures, 
        tbl_revisit, 
        tbl_readmit
    ) %>%
    select(Event, pharmacy, traditional) %>%
    mutate(p = p) %>%
    knitr::kable(
        format = "latex",
        digits = c(0, 1, 1, 2), 
        booktabs = TRUE,
        col.names = c("Event", "%", "%", "p-value"), 
        caption = paste(
            "Outcomes and Events During Warfarin Therapy.",
            "Hemoglobin decrease includes a drop of >/= 2 g/dL within a 48-hour period.",
            "A new Vizient metric evaluates the percent of patients on warfarin for more than two days who experience an INR >/= 5 at any point during the hospitalization after warfarin initation.", 
            "Reversal medications include phytonadione, 4-factor PCC, factor VIIa, aminocaproic acid, and tranexamic acid", 
            "Unplanned return includes any ER visit, observation, or inpatient stay at any MH hospital within 30 days of discharge from MH-TMC",
            sep = " "
        )
    ) %>%
    add_header_above(c("", "Pharmacy", "Traditional", ""), align = "r", line = FALSE)
```

# Conclusion

The Pharmacy Dosing Service continued to provide safe and effective management of warfarin for patients in fiscal year 2018. 
