---
title: "Warfarin Pharmacy Dosing Service analysis for Fiscal Year 2020"
author: "Brian Gulbis, PharmD, BCPS"
date: "November 9, 2020"
output:
  tufte::tufte_handout:
    latex_engine: xelatex
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(
    tidy = FALSE, 
    echo = FALSE, 
    message = FALSE, 
    warning = FALSE, 
    cache.extra = packageVersion('tufte'), 
    kable.force.latex = TRUE
)

options(htmltools.dir.version = FALSE)
```

```{r data, message=FALSE}
library(data.table)
library(tidyverse)
library(lubridate)
library(broom)
library(mbohelpr)
library(themebg)
library(kableExtra)

cur_fy <- 2020

data_dir <- "../data/tidy/fy2020"

find_ind <- function(x, reg) {
    stringr::str_detect(x, regex(reg, ignore_case = TRUE))
}

data_blood <- get_data(data_dir, "blood_products") 

data_consult_orders <- get_data(data_dir, "consult_orders") %>%
    filter(str_detect(nurse_unit, "^HH|^HVI")) %>%
    mutate(
        order_month = floor_date(order_datetime, unit = "month"),
        fiscal_year = year(order_month %m+% months(6))
    )
    
data_consult_tasks <- get_data(data_dir, "consult_tasks") %>%
    filter(
        str_detect(nurse_unit, "^HH|^HVI"),
        task_date < ymd("2020-07-01")
    ) %>%
    mutate(
        task_month = floor_date(task_date, unit = "month"),
        fiscal_year = year(task_month %m+% months(6))
    )

data_demographics <- get_data(data_dir, "demographics")

data_doac_doses <- get_data(data_dir, "doac_doses") %>%
    filter(str_detect(nurse_unit, "^HH|^HVI")) %>%
    mutate(
        med_month = floor_date(med_datetime, unit = "month"),
        fiscal_year = year(med_month %m+% months(6))
    )
    
data_labs <- get_data(data_dir, "labs")
data_measures <- get_data(data_dir, "measures")
data_reencounters <- get_data(data_dir, "reencounters")
data_reversal_meds <- get_data(data_dir, "reversal_meds")
data_warfarin_details <- get_data(data_dir, "warfarin_details")

data_warfarin_doses <- get_data(data_dir, "warfarin_doses") %>%
    filter(str_detect(nurse_unit, "^HH|^HVI")) %>%
    mutate(
        med_day = floor_date(med_datetime, unit = "day"),
        med_month = floor_date(med_datetime, unit = "month"),
        fiscal_year = year(med_month %m+% months(6))
    )

data_warfarin_home_meds <- get_data(data_dir, "warfarin_home_meds")
data_warfarin_orders <- get_data(data_dir, "warfarin_orders")

df_indications <- data_warfarin_details %>%
    filter(detail == "Warfarin Indication") %>%
    mutate(
        indication = case_when(
            find_ind(
                result_value, 
                "Atrial fibrillation|a(.*)?fib|a(.*)?flutter"
            ) ~ "afib",
            find_ind(
                result_value, 
                "Deep vein thrombosis|DVT(?!( prophylaxis))|VTE"
            ) ~ "dvt",
            find_ind(
                result_value, 
                "Pulmonary embolism|PE"
            ) ~ "pe",
            find_ind(
                result_value, 
                "Heart valve \\(Mech/porc/bioprost\\)|valve|avr|mvr"
            ) ~ "valve",
            find_ind(
                result_value, 
                "st(ro|or)ke|cva|ica|mca"
            ) ~ "stroke",
            find_ind(
                result_value, 
                "vad|hm[ ]?ii|heart( )?mate|heartware|syncardia|total artificial heart|tah"
            ) ~ "vad",
            find_ind(
                result_value, 
                "throm|clot|emboli|occl"
            ) ~ "thrombus",
            find_ind(
                result_value, 
                "malig|anti(.)?phos|lupus|apla|hypercoag|deficien|leiden|fvl|factor v"
            ) ~ "hypercoag",
            find_ind(
                result_value, 
                "prophylax"
            ) ~ "prophylax",
            TRUE ~ "other"
        )
    ) %>%
    distinct(encounter_id, indication) 
    # mutate(value = TRUE) %>%
    # spread(indication, value, fill = FALSE)
        
df_inr_goal <- data_warfarin_details %>%
    filter(
        detail == "INR Range",
        result_value != ""
    ) %>%
    mutate(inr_range = result_value) %>%
    mutate_at(
        "inr_range", 
        str_replace_all, 
        pattern = " |", 
        replacement = ""
    ) %>%
    mutate_at(
        "inr_range", 
        str_replace_all, 
        pattern = regex("to|--|/|=|,|-\\.", ignore_case = TRUE), 
        replacement = "-"
    ) %>%
     mutate_at(
        "inr_range",
        str_replace_all,
        pattern = ">1.5-<2.5",
        replacement = "1.5-2.5"
    ) %>%
    mutate_at(
        "inr_range",
        str_replace_all,
        pattern = regex(
            c(
                # "^1.[5-9]$" = "1.5-2",
                "^2$|^2.0$" = "1.5-2.5",
                "^2\\.[1-4]$" = "2-2.5",
                "^2.5$" = "2-3",
                # "^2.[6-9]$" = "2.5-3",
                "^3$" = "2.5-3.5",
                # "^3.5$" = "3-4",
                # "^22" = "2",
                # "2.53.5" = "2.5-3.5",
                "^23$" = "2-3",
                "30$" = "3",
                "^1-5" = "1.5"
            ),
            ignore_case = TRUE
        )
    ) %>%
    extract(
        inr_range,
        c("goal_low", "goal_high"), 
        regex = "([0-9\\.]+ ?)-( ?[0-9\\.]+)",
        remove = FALSE, 
        convert = TRUE
    ) %>%
    filter(
        !is.na(goal_low),
        goal_low < goal_high
    ) %>%
    mutate_at(c("goal_low", "goal_high"), as.numeric) %>%
    select(-result_value, -inr_range)

```

```{r tidying}
df_consult_pts <- data_consult_orders %>%
    arrange(encounter_id, order_datetime) %>%
    distinct(encounter_id, .keep_all = TRUE) %>%
    mutate(
        consult = TRUE,
        consult_date = floor_date(order_datetime, unit = "day")
    ) %>%
    select(
        encounter_id, 
        consult_date,
        nurse_unit_consult = nurse_unit,
        med_service, 
        consult
    )

df_warf_start <- data_warfarin_doses %>%
    arrange(encounter_id, med_datetime) %>%
    distinct(encounter_id, .keep_all = TRUE) %>%
    mutate(warfarin_start = floor_date(med_datetime, unit = "day")) %>%
    select(
        encounter_id,
        med_month,
        warfarin_start, 
        nurse_unit_warfarin = nurse_unit, 
        fiscal_year
    )

df_demog <- data_demographics %>%
    inner_join(df_warf_start, by = "encounter_id") %>%
    left_join(df_consult_pts, by = "encounter_id") %>%
    mutate_at("consult", list(~coalesce(., FALSE))) %>%
    mutate(
        consult_day = difftime(
            consult_date, 
            warfarin_start, 
            units = "days"
        )
    )

df_warf_month <- data_warfarin_doses %>%
    count(med_month, name = "warfarin") %>%
    mutate_at("med_month", as_date)

df_consult_month <- data_consult_tasks %>%
     count(task_month, name = "consults")

df_doac_month <- data_doac_doses %>%
    mutate(med_day = floor_date(med_datetime, unit="day")) %>%
    distinct(encounter_id, medication, med_month, med_day) %>%
    count(med_month, name = "doac") %>%
    mutate_at("med_month", as_date)

df_warf_ord <- data_warfarin_orders %>%
    rename(
        nurse_unit_order = nurse_unit,
        med_service_order = med_service
    ) %>%
    arrange(encounter_id, order_datetime) %>%
    mutate(order_date = floor_date(order_datetime, unit = "day")) %>%
    inner_join(df_warf_start, by = c("encounter_id", "order_date" = "warfarin_start")) %>%
    left_join(df_consult_pts, by = "encounter_id") %>%
    mutate_at("consult", list(~coalesce(., FALSE)))

df_warf_duration <- data_warfarin_doses %>%
    arrange(encounter_id, med_datetime) %>%
    group_by(encounter_id) %>%
    summarize_at("med_day", list(warfarin_start = first, warfarin_stop = last))

 df_doses_daily <- data_warfarin_doses %>%
    arrange(encounter_id, med_datetime) %>%
    group_by(encounter_id, fiscal_year, med_day) %>%
    summarize_at("dose", sum, na.rm = TRUE) %>%
    group_by(encounter_id, fiscal_year) %>%
    mutate(warf_day = difftime(med_day, first(med_day), units = "days") + 1) %>%
    mutate_at("warf_day", as.integer) %>%
    filter(warf_day <= 10) %>%
    ungroup() %>%
    select(encounter_id, fiscal_year, warf_day, dose) 

df_inr_daily <- df_warf_duration %>%
    inner_join(data_labs, by = "encounter_id") %>%
    filter(lab == "INR") %>%
    mutate(
        lab_day = floor_date(lab_datetime, unit = "days"),
        lab_month = floor_date(lab_datetime, unit = "month"),
        warf_day = difftime(lab_day, warfarin_start, units = "days") + 1,
        fiscal_year = year(lab_month %m+% months(6))
    ) %>%
    mutate_at("warf_day", as.integer) %>%
    filter(
        lab_day >= warfarin_start,
        lab_day <= warfarin_stop,
        warf_day <= 10
    ) %>%
    group_by(encounter_id, fiscal_year, lab_day, warf_day) %>%
    summarize_at("result_value", max, na.rm = TRUE) %>%
    ungroup() %>%
    select(encounter_id, fiscal_year, warf_day, inr = result_value) 

df_consult_daily <- df_warf_duration %>%
    mutate_at(c("warfarin_start", "warfarin_stop"), as_date) %>%
    inner_join(data_consult_tasks, by = "encounter_id") %>%
    mutate(
        warf_day = difftime(task_date, warfarin_start, units = "days") + 1,
        task_month = floor_date(task_date, unit = "month"),
        fiscal_year = year(task_month %m+% months(6))
    ) %>%
    mutate_at("warf_day", as.integer) %>%
    filter(
        task_date >= warfarin_start,
        task_date <= warfarin_stop,
        warf_day <= 10
    ) %>%
    select(encounter_id, fiscal_year, warf_day) %>%
    mutate(consult = TRUE)

df_doses_inr <- df_doses_daily %>%
    full_join(
        df_inr_daily, 
        by = c("encounter_id", "fiscal_year", "warf_day")
    ) %>%
    full_join(
        df_consult_daily,
        by = c("encounter_id", "fiscal_year", "warf_day")
    ) %>%
    arrange(encounter_id, warf_day)

df_group <- df_doses_inr %>%
    filter(warf_day < 3) %>%
    group_by(encounter_id, fiscal_year) %>%
    summarize_at("consult", sum, na.rm = TRUE) %>%
    mutate_at("consult", list(~. > 0))

df_doses_cnt <- df_doses_inr %>%
    group_by(encounter_id, fiscal_year) %>%
    summarize(
        doses_n = sum(!is.na(dose), na.rm = TRUE)
    )

df_doses_3 <- df_doses_cnt %>%
    inner_join(df_group, by = c("encounter_id", "fiscal_year")) %>%
    filter(
        doses_n >= 3,
        fiscal_year == cur_fy
    )
  
df_goal_eq2_3 <- df_inr_goal %>%
    semi_join(df_doses_3, by = "encounter_id") %>%
    filter(
        goal_low == 2,
        goal_high == 3
    ) %>%
    distinct(encounter_id) %>%
    mutate(goal2_3 = TRUE)

tidy_pts <- df_doses_3 %>%
    left_join(df_goal_eq2_3, by = "encounter_id") %>%
    mutate_at("goal2_3", list(~coalesce(., FALSE))) %>%
    ungroup()

```

```{r pct_pts}
df_cnt <- df_demog %>%
    add_count(fiscal_year, name = "n_fy") %>%
    count(fiscal_year, n_fy, consult) %>%
    mutate(pct = (n / n_fy) * 100)

n_consult_cur_fy <- df_cnt %>%
    filter(
        fiscal_year == cur_fy,
        consult
    )

pct_warf_use <- df_cnt %>%
    distinct(fiscal_year, n_fy) %>%
    pivot_wider(names_from = fiscal_year, values_from = n_fy) %>%
    mutate(pct_chg = (`2018` - `2020`) / `2018` * 100)
```


# Summary

The Department of Pharmacy's Warfarin Dosing Service was consulted on `r round(n_consult_cur_fy$pct, 0)`% of warfarin patients in fiscal year 2020 (FY20). The service lines with the highest utilization were Cardiology, Internal Medicine, CV Surgery, and Neurology. Patients managed by the pharmacy service were more likely to reach the target therapeutic range compared with traditional provider management of warfarin. 

# Warfarin Utilization

There was a `r round(pct_warf_use$pct_chg, 0)`% decrease in the total number of warfarin doses administered in FY20 compared with FY18. This was due to a combination of reduced volume due to COVID-19, as well as a continued increase in the use of Direct Oral Anticoagulants (DOACs), such as apixaban and rivaroxaban (see Figure 1). The Pharmacy Dosing Service was consulted on `r round(n_consult_cur_fy$pct, 0)`% of warfarin patients in FY20.

```{r, fig.cap="The number of monthly doses of warfarin managed by the Pharmacy Dosing Service, by traditional providers, and the number of monthly doses of direct oral anticoagulants (DOACs) over the past three fiscal years."}

df_warf_month %>%
    inner_join(df_consult_month, by = c("med_month" = "task_month")) %>%
    inner_join(df_doac_month, by = "med_month") %>%
    mutate(traditional = warfarin - consults) %>%
    select(-warfarin) %>%
    gather(key, value, -med_month) %>%
    mutate(
        across(key, as_factor),
        across(key, fct_rev)
    ) %>%
    ggplot(aes(x = med_month, y = value, color = key, size = key)) +
    annotate(
        "text", 
        x = mdy("4/15/2017"), # 7/15
        y = 525, # 1175
        label = "Providers", 
        family = "serif", 
        color = "#4daf4a",
        size = 2.5
    ) +
    annotate(
        "text", 
        x = mdy("4/15/2017"), # 7/20
        y = 625, # 715
        label = "Pharmacy", 
        family = "serif", 
        color = "#377eb8",
        size = 2.5
    ) +
    annotate(
        "text", 
        x = mdy("4/15/2017"), # 7/1
        y = 450, # 500
        label = "DOACs", 
        family = "serif", 
        color = "#ff7f00",
        size = 2.5
    ) +
    geom_line() +
    ggtitle("Monthly doses of oral anticoagulants") +
    scale_x_date(NULL, date_breaks = "6 months", date_labels = "%b %y", limits = c(as.Date("2017-03-15"), NA)) +
    scale_y_continuous("Doses") +
    scale_color_manual(NULL, values = c("#4daf4a", "#ff7f00", "#377eb8")) + # c("#377eb8", "#ff7f00", "#4daf4a")
    scale_size_manual(NULL, values = c(0.5, 0.5, 1)) +
    coord_cartesian(ylim = c(0, 1250)) +
    theme_bg(base_family = "serif") +
    theme(
        legend.position = "none", 
        plot.title = element_text(size = 10), 
        axis.title = element_text(size = 9), 
        axis.text = element_text(size = 8)
    )

```

The highest volume of warfarin use occurred in patients on the Cardiology service line, followed by Internal Medicine, CV Surgery, and Neurology (see Figure 2). The Cardiology service line primarily utilizes traditional provider management of warfarin, although Clinical Pharmacists and Clinical Pharmacist-Specialists participate in the dosing of warfarin on these patients outside of a formal consult to the Pharmacy Dosing Service.

```{r, fig.cap="The number of warfarin patients by the primary service at the time of warfarin initiation. The Cardiology service line remains the largest user of warfarin. The Pharmacy Dosing Service manages the majority of warfarin dosing for patients on most other service lines."}
df_warf_ord %>%
    filter(fiscal_year == cur_fy) %>%
    mutate_at(
        "med_service_order",
        str_replace_all,
        pattern = c(
            "Thoracic/Cardiac Sur Service" = "CV Surgery",
            "Therapy - Phys/Occ/Speech/Cardiac" = "Rehab",
            "Pulmonology/Respiratory Therapy" = "Pulmonary"
        )
    ) %>%
    mutate_at("med_service_order", fct_infreq) %>%
    mutate_at("med_service_order", fct_lump, n = 7) %>%
    mutate_at("med_service_order", fct_rev) %>%
    ggplot(aes(x = med_service_order, fill = consult)) +
    geom_bar() +
    ggtitle("Warfarin utilization by primary service") +
    xlab(NULL) +
    ylab("Patients") +
    scale_fill_manual(
        NULL, 
        values = c("#4daf4a", "#377eb8"), 
        labels = c("Traditional", "Pharmacy")
    ) +
    coord_flip(ylim = c(0, 1100)) +
    theme_bg_print(base_family = "serif") +
    theme(
        legend.position = "top",
        plot.title = element_text(size = 10), 
        axis.title = element_text(size = 9), 
        axis.text = element_text(size = 8),
        legend.text = element_text(size = 8),
        legend.box.margin = margin(t = 0, b = 0, unit='cm')
    ) + 
    guides(fill = guide_legend(override.aes = list(size = 1)))
```

The warfarin dosing for patients with a ventricular assist device (VAD) is primarily managed by traditional provider dosing. The Pharmacy Dosing Service manages the majority of patients with all other indications for anticoagulation (see Figure 3). 

```{r, fig.cap="Indications for oral anticoagulation. The most common indication for warfarin was atrial fibrillation (A.fib), followed by ventricular assist devices (VADs), mechanical heart valves, deep vein thrombosis (DVTs), and pulmonary embolisms (PEs)."}
df_demog %>%
    filter(fiscal_year == cur_fy) %>%
    inner_join(df_indications, by = "encounter_id") %>%
    mutate_at(
        "indication",
        str_replace_all,
        pattern = c(
            "afib" = "A.fib",
            "vad" = "VAD",
            "valve" = "Valve",
            "dvt" = "DVT",
            "hypercoag" = "Hypercoaguable",
            "^pe" = "PE",
            "thrombus" = "Thrombus",
            "stroke" = "Stroke",
            "other" = "Other"
        )
    ) %>% 
    mutate_at("indication", fct_infreq) %>%
    mutate_at("indication", fct_rev) %>%
    # mutate_at("fiscal_year", as_factor) %>%
    # add_count(consult, name = "n_grp") %>%
    # count(consult, n_grp, indication) %>%
    # mutate(pct = n / n_grp) %>%
    ggplot(aes(x = indication, fill = consult)) +
    geom_bar() +
    ggtitle("Warfarin indications") +
    xlab(NULL) +
    ylab("Patients") + # calculate percent?
    scale_fill_manual(
        NULL, 
        values = c("#4daf4a", "#377eb8"), 
        labels = c("Traditional", "Pharmacy")
    ) +
    coord_flip() +
    theme_bg_print(base_family = "serif") +
    theme(
        legend.position = "top",
        plot.title = element_text(size = 10), 
        axis.title = element_text(size = 9), 
        axis.text = element_text(size = 8),
        legend.text = element_text(size = 8),
        legend.box.margin = margin(t = 0, b = 0, unit='cm')
    ) + 
    guides(fill = guide_legend(override.aes = list(size = 1)))
```

# Dosing and INR Response

```{r n_pts}
n_pts <- tidy_pts %>%
    count(consult) %>%
    mutate_at("consult", list(~if_else(., "Pharmacy", "Traditional"))) %>%
    pivot_wider(names_from = consult, values_from = n)

```

Patients managed by the Pharmacy Dosing Service typically received a higher daily warfarin dose than those managed by other providers. Only patients with an INR target between 2 and 3 were included to ensure the goal therapeutic range was equivalent in both groups.

```{r, fig.cap="The daily warfarin dose administered to patients in each group. Patients in the Pharmacy Dosing Service group received a higher daily warfarin dose than those in the traditional provider management group. The p-value is < 0.01 for each day of warfarin therapy."}
p <- df_doses_inr %>%
    semi_join(tidy_pts, by = "encounter_id") %>%
    filter(
        warf_day <= 10,
        fiscal_year == cur_fy
    ) %>%
    mutate_at("consult", list(~coalesce(., FALSE))) %>%
    group_by(warf_day) %>%
    do(tidy(wilcox.test(dose ~ consult, data = .))) %>%
    mutate_at("p.value", round, digits = 3)

df_doses_inr %>%
    semi_join(tidy_pts, by = "encounter_id") %>%
    filter(
        warf_day <= 10,
        fiscal_year == cur_fy
    ) %>%
    mutate_at("consult", list(~coalesce(., FALSE))) %>%
    ggplot(aes(x = warf_day, y = dose, color = consult)) +
    annotate(
        "text", 
        x = 1.4, 
        y = 6.2, 
        label = "Pharmacy",
        color = "#377eb8",
        family = "serif",
        size = 2.5
    ) + 
    annotate(
        "text", 
        x = 1.4, 
        y = 4.45, 
        label = "Traditional",
        color = "#4daf4a", 
        family = "serif",
        size = 2.5
    ) +
    geom_smooth() +
    ggtitle("Warfarin daily doses") +
    scale_x_continuous("Day of therapy", breaks = seq(1, 10, 1)) +
    ylab("Dose (mg)") +
    scale_color_manual(
        NULL,
        values = c("#4daf4a", "#377eb8"),
        labels = c("Traditional", "Pharmacy")
    ) + 
    coord_cartesian(ylim = c(0, 10)) +
    theme_bg_print(base_family = "serif") +
    theme(
        legend.position = "none",
        plot.title = element_text(size = 10), 
        axis.title = element_text(size = 9), 
        axis.text = element_text(size = 8)
    )
  
    # mutate_at("warf_day", as_factor) %>%
    # ggplot(aes(x = warf_day, y = dose, fill = consult, color = consult)) +
    # geom_boxplot(outlier.shape = NA) +
    # xlab("Day of therapy") +
    # ylab("Dose (mg)") +
    # scale_color_manual(
    #     NULL,
    #     values = c("grey25", "grey50"),
    #     labels = c("Traditional", "Pharmacy")
    # ) + 
    # scale_fill_manual(
    #     NULL, 
    #     values = c("grey50", "black"), 
    #     labels = c("Traditional", "Pharmacy")
    # ) +
    # coord_cartesian(ylim = c(0, 15)) +
    # theme_bg_print(base_family = "serif") +
    # theme(legend.position = "top")

```

Despite starting with a lower baseline INR, patients managed by the Pharmacy Dosing Service had a greater rise in INR over the first 10 days of therapy, and were more likely to reach an INR around 2 on day 5 (see Figure 5). 

```{r, fig.cap="The daily INR value for patients in each group over the first 10 days of warfarin therapy. There was a larger increase in the INR for patients managed by the Pharmacy Dosing Service. The p-value was < 0.05 on all days except days 5 and 6 where there was no significant difference."}

p <- df_doses_inr %>%
    semi_join(tidy_pts, by = "encounter_id") %>%
    # semi_join(df_goal_eq2_3, by = "encounter_id") %>%
    filter(
        # warf_day <= 7
        fiscal_year == cur_fy,
        !is.na(inr)
    ) %>%
    mutate_at("consult", list(~coalesce(., FALSE))) %>%
    mutate_at("warf_day", round, digits = 0) %>%
    group_by(warf_day) %>%
    do(tidy(wilcox.test(inr ~ consult, data = .))) %>%
    mutate_at("p.value", round, digits = 3)

df_doses_inr %>%
    semi_join(tidy_pts, by = "encounter_id") %>%
    # semi_join(df_goal_eq2_3, by = "encounter_id") %>%
    filter(
        # warf_day <= 7
        fiscal_year == cur_fy,
        !is.na(inr)
    ) %>%
    mutate_at("consult", list(~coalesce(., FALSE))) %>%
    ggplot(aes(x = warf_day, y = inr, color = consult)) +
    geom_hline(yintercept = 2, linetype = "dashed", color = "grey75") +
    annotate(
        "text", 
        x = 1.5, 
        y = 1.35, 
        label = "Pharmacy",
        color = "#377eb8",
        family = "serif",
        size = 2.5
    ) + 
    annotate(
        "text", 
        x = 1.5, 
        y = 1.85, 
        label = "Traditional",
        color = "#4daf4a", 
        family = "serif",
        size = 2.5
    ) +
    geom_smooth() +
    ggtitle("INR response") +
    scale_x_continuous("Day of therapy", breaks = seq(1, 10, 1)) +
    ylab("INR") +
    scale_color_manual(
        NULL,
        values = c("#4daf4a", "#377eb8"),
        labels = c("Traditional", "Pharmacy")
    ) + 
    coord_cartesian(ylim = c(1, 3)) +
    theme_bg_print(base_family = "serif") +
    theme(
        legend.position = "none",
        plot.title = element_text(size = 10), 
        axis.title = element_text(size = 9), 
        axis.text = element_text(size = 8)
    )

```

# Outcomes and Events

```{r high_inr}
df_high_inr <- df_doses_inr %>%
    filter(
        inr >= 5
    ) %>%
    distinct(encounter_id) %>%
    mutate(high_inr = TRUE)
```

```{r hgb_drop}
df_hgb <- df_warf_duration %>%
    inner_join(data_labs, by = "encounter_id") %>%
    arrange(encounter_id, lab_datetime) %>%
    filter(lab == "Hgb") %>%
    mutate(
        lab_day = floor_date(lab_datetime, unit = "days"),
        lab_month = floor_date(lab_datetime, unit = "month"),
        warf_day = difftime(lab_day, warfarin_start, units = "days") + 1,
        fiscal_year = year(lab_month %m+% months(6))
    ) %>%
    mutate_at("warf_day", as.integer) %>%
    filter(
        lab_day >= warfarin_start,
        lab_day <= warfarin_stop + 1,
    ) 

dt_hgb <- as.data.table(df_hgb)

dt_window <- dt_hgb[, .(
    encounter_id, 
    window_low = lab_datetime - hours(48), 
    window_high = lab_datetime
)]

dt_hgb_max <- dt_hgb[
    dt_window, 
    on=.(
        encounter_id, 
        lab_datetime >= window_low, 
        lab_datetime <= window_high
    ), 
    .(max_hgb = max(result_value)), 
    by=.EACHI
    ][, lab_datetime := NULL
      ][, head(.SD, 1), by = c("encounter_id", "lab_datetime")]

dt_hgb_drop <- dt_hgb[dt_hgb_max, on=.(encounter_id, lab_datetime)]

df_hgb_drop <- as_tibble(dt_hgb_drop) %>%
    arrange(encounter_id, lab_datetime) %>%
    mutate(hgb_chg = result_value - max_hgb) %>%
    filter(
        hgb_chg <= -2,
        warf_day <= 10
    ) %>% 
    distinct(encounter_id) %>%
    mutate(hgb_drop = TRUE)

```

```{r prbc}
df_prbc <- data_blood %>%
    inner_join(df_warf_duration, by = "encounter_id") %>%
    filter(
        str_detect(product, regex("rbc", ignore_case = TRUE)),
        event_datetime >= warfarin_start,
        event_datetime <= warfarin_stop 
    ) %>%
    distinct(encounter_id) %>%
    mutate(prbc = TRUE)

```

```{r ffp}
df_ffp <- data_blood %>%
    inner_join(df_warf_duration, by = "encounter_id") %>%
    filter(
        str_detect(product, regex("ffp|plasma", ignore_case = TRUE)),
        event_datetime >= warfarin_start,
        event_datetime <= warfarin_stop 
    ) %>%
    distinct(encounter_id) %>%
    mutate(ffp = TRUE)

```

```{r reversal}
df_reversal <- data_reversal_meds %>%
    inner_join(df_warf_duration, by = "encounter_id") %>%
    filter(
        med_datetime >= warfarin_start,
        med_datetime <= warfarin_stop 
    ) %>%
    distinct(encounter_id) %>%
    mutate(reversal_med = TRUE)
```

```{r revisits}
df_dispo <- df_demog %>%
    select(encounter_id, disch_disposition) %>%
    # inner_join(tidy_pts, by = "encounter_id") %>%
    mutate(
        dispo_group = case_when(
            str_detect(disch_disposition, "Hospice|Deceased") ~ "Deceased",
            str_detect(disch_disposition, "Home|Left") ~ "Home",
            str_detect(disch_disposition, "Facility|REHAB|Care|DC/TF") ~ "Transfer"
        )
    )

df_revisit <- df_dispo %>%
    filter(dispo_group == "Home") %>%
    inner_join(data_reencounters, by = "encounter_id") %>%
    distinct(encounter_id) %>%
    mutate(revisit = TRUE)
```

```{r readmits}
df_readmit <- df_dispo %>%
    filter(dispo_group == "Home") %>%
    inner_join(data_reencounters, by = "encounter_id") %>%
    filter(readmit_type == "Inpatient") %>%
    distinct(encounter_id) %>%
    mutate(readmit = TRUE)
```

```{r}
df_outcomes <- tidy_pts %>%
    left_join(df_dispo, by = "encounter_id") %>%
    left_join(df_high_inr, by = "encounter_id") %>%
    left_join(df_hgb_drop, by = "encounter_id") %>%
    left_join(df_prbc, by = "encounter_id") %>%
    left_join(df_ffp, by = "encounter_id") %>%
    left_join(df_reversal, by = "encounter_id") %>%
    left_join(df_revisit, by = "encounter_id") %>%
    left_join(df_readmit, by = "encounter_id") %>%               
    # filter(fiscal_year == cur_fy) %>%
    select(encounter_id, consult, dispo_group, high_inr:readmit) %>%
    mutate_at(
        c(
            "high_inr", 
            "hgb_drop", 
            "prbc",
            "ffp",
            "reversal_med"
        ),
        list(~coalesce(., FALSE))
    ) %>%
    mutate_at(
        c("revisit", "readmit"),
        list(~if_else(dispo_group == "Home", coalesce(., FALSE), .))
    ) %>%
    mutate_at("consult", list(~if_else(., "Pharmacy", "Traditional"))) %>%
    add_count(consult, name = "group_n")

df_outcomes_cnt <- df_outcomes %>%
    group_by(consult, group_n) %>%
    summarize_at(
        c(
            "high_inr", 
            "hgb_drop", 
            "prbc",
            "ffp",
            "reversal_med",
            "revisit", 
            "readmit"
        ),
        sum,
        na.rm = TRUE
    )

```

```{r eval=FALSE}
df_p <- df_outcomes %>%
    mutate_at(c("consult", "dispo_group"), as_factor) 

chisq.test(x = df_p$consult, y = df_p$dispo_group)
chisq.test(x = df_p$consult, y = df_p$high_inr)
chisq.test(x = df_p$consult, y = df_p$hgb_drop)
chisq.test(x = df_p$consult, y = df_p$prbc)
chisq.test(x = df_p$consult, y = df_p$ffp)
chisq.test(x = df_p$consult, y = df_p$reversal_med)
chisq.test(x = df_p$consult, y = df_p$revisit)
chisq.test(x = df_p$consult, y = df_p$readmit)

```

There were no differences in supratherapeutic INR, bleeding, need for transfusion, or need for anticoagulation reversal between the two groups (see Table 1).

```{r outcomes}
df_outcomes_pct <- df_outcomes %>%
    select(-group_n) %>%
    group_by(consult) %>%
    summarize_at(
        c(
            "high_inr", 
            "hgb_drop", 
            "prbc",
            "ffp",
            "reversal_med",
            "revisit",
            "readmit"
        ),
        mean,
        na.rm = TRUE
    ) %>%
    pivot_longer(
        high_inr:readmit, 
        names_to = "outcome", 
        values_to = "pct"
    ) %>%
    mutate_at(
        "outcome",
        str_replace_all,
        pattern = c(
            "high_inr" = "High INR",
            "hgb_drop" = "Hemoglobin drop",
            "prbc" = "Transfuse PRBC",
            "ffp" = "Transfuse FFP",
            "reversal_med" = "Adminster reversal agent",
            "revisit" = "Unplanned return*",
            "readmit" = "Hospital readmission*"
        )
    ) %>%
    mutate_at("pct", list(~. * 100)) %>%
    pivot_wider(
        names_from = "consult",
        values_from = "pct"
    )

df_outcomes_pct %>%
    filter(
        outcome != "Unplanned return*",
        outcome != "Hospital readmission*"
    ) %>%
    knitr::kable(
        format = "latex",
        digits = 0,
        booktabs = TRUE,
        col.names = c(
            "Outcome",
            "Pharmacy (%)",
            "Traditional (%)"
        ),
        caption = "Outcomes and events during the first 10 days of warfarin therapy."
    ) %>%
    footnote(
        number = c(
            "High INR is defined as an INR of 5 or greater after warfarin initiation.", 
            "Hemoglobin decrease includes a drop of >/= 2 g/dL within a 48-hour period.", 
            "Reversal medications include phytonadione, 4-factor PCC, factor VIIa, aminocaproic acid, and tranexamic acid."
        ),
        threeparttable = TRUE,
        general_title = "",
        fixed_small_size = TRUE
    )

```

A larger percent of patients managed by the Pharmacy Dosing Service were discharged to another facility for additional care (see Table 2). Among patients who were discharged home, a larger percent of patients in the traditional provider group had an unplanned return or readmission within 30 days.

```{r dispo}
df_outcome_dispo <- df_outcomes %>%
    filter(!is.na(dispo_group)) %>%
    count(consult, group_n, dispo_group) %>%
    mutate(dispo_pct = n / group_n * 100) %>%
    select(-group_n, -n) %>%
    pivot_wider(
        names_from = "consult",
        values_from = "dispo_pct"
    ) %>%
    rename(outcome = dispo_group)

df_outcome_revisit <- df_outcomes_pct %>%
    filter(
        outcome %in% c("Unplanned return*", "Hospital readmission*")
    )

df_outcome_dispo %>%
    bind_rows(df_outcome_revisit) %>%
    knitr::kable(
        format = "latex",
        digits = 0,
        booktabs = TRUE,
        col.names = c(
            "Outcome",
            "Pharmacy (%)",
            "Traditional (%)"
        ),
        caption = "Discharge disposition and unplanned return within 30 days."
    ) %>%
    pack_rows(
        "Discharge disposition*", 
        start_row = 1, 
        end_row = 3,
        bold = FALSE) %>%
    footnote(
        symbol = "p-value < 0.001",
        number = "Unplanned return includes any ER visit, observation, or inpatient stay at any MH hospital within 30 days of discharge among patients who were discharged home.",
        footnote_order = c("symbol", "number"),
        threeparttable = TRUE,
        general_title = "",
        fixed_small_size = TRUE
    )
```

# Conclusion

The Pharmacy Dosing Service continued to provide safe and effective management of warfarin for patients in fiscal year 2019. 
